---
  title: 'Data Analysis 3: Week 6'
author: "Alexey Bessudnov"
date: "21 February 2019"
output: github_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)
```
Plan for today:
  
- Assignment 2: solution
- Assignment 3
- Simple data visualisation 
- Homework for next week

## Exercises

1. Read the *indresp* file from Wave 8 and keep the following variables: pidp, derived sex and age, ethnic group (h_indresp), government office region (h_gor_dv), and net personal income (h_fimnnet_dv).

For all charts use ggplot2. You may need to clean and recode variable before visualising.

We will start with univariate distributions.

```{r}
library(tidyverse)
library(ggplot2)

w8 <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")

w8 <- w8 %>% 
  select(pidp, h_sex_dv, h_age_dv, h_racel_dv, h_gor_dv, h_fimnnet_dv)
```

2. Visualise the distribution of income with a histogram, a density plot and a box plot.

```{r}
summary(w8$h_fimnnet_dv) # to help decide the limits (xlim or ylim)

w8 %>% 
  ggplot(aes(x = h_fimnnet_dv)) +
  geom_histogram(binwidth = 100) +
  xlab("Net Monthly Personal Income") +
  xlim(-500, 7000)

w8 %>% 
  ggplot(aes(x = h_fimnnet_dv)) +
  geom_density() +
  xlim(-500, 7000)
  
w8 %>% 
  ggplot(aes(y = h_fimnnet_dv)) +
  geom_boxplot() +
  ylim(-500, 5000)
```

3. Visualise the distribution of sex with a bar graph. 

Bivariate distributions.

```{r}
w8 %>% 
  mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male", 
                           ifelse(h_sex_dv == 2, "female", NA))) %>% 
  filter(!is.na(h_sex_dv)) %>% 
  ggplot(aes(x = h_sex_dv)) +
  geom_bar() +
  xlab("Sex")
  
## Alternatively, to show proportions instead of counts

w8 %>% 
  mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male", 
                           ifelse(h_sex_dv == 2, "female", NA))) %>% 
  filter(!is.na(h_sex_dv)) %>% 
  ggplot(aes(x = h_sex_dv)) +
  geom_bar(aes(y = (..count..)/ sum(..count..) * 100)) +
  xlab("Sex") +
  ylab("Percentage") +
  scale_y_continuous(breaks = seq(0, 60, 10))
```

4. Create a bar chart showing mean income by region. Then make a grouped bar chart showing mean income by region for men and women separately.

```{r}
income <- w8 %>% 
  mutate(region = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire and the Humber",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>% 
  filter(!is.na(region)) %>% 
  group_by(region) %>% 
  summarise(
    meanIncome = mean(h_fimnnet_dv, na.rm = TRUE),
    medianIncome = median(h_fimnnet_dv, na.rm = TRUE)
  )

income %>% 
  ggplot(aes(y = medianIncome, x = region)) +
  geom_bar(stat = "identity")+
  xlab("Region") +
  ylab("Mean Net Monthly Personal Income") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

5. Make a dot plot showing the same information as above (without splitting by sex). Sort regions in the descending order by mean income.

6. Make a line chart showing mean income by age. Make another chart showing the association for men and women separately.

7. Create a scatterplot showing the association between age and income for randomly selected 100 men in the sample. Now display a smoothing curve on the same graph showing the association between age and income.