---
title: "dplyr notes"
author: "Youran Xu"
date: "February 1st, 2019"

---

# dplyr

```{r}
library(nycflights13)
library(tidyverse)
```

All dplyr basics work in a similar way:
1. 1st arguement: a data frame
2. 2nd: commands in relation to variables
3. result: a new data frame

## `fliter()`

- (dataset, logical expressions to fliter)
- dplyr doesn't modify inputs, so always needs assignments
- `is.na(x)` to determine whether a value is missing
    - to include NA values: `filter(data, is.na(x) | x > 1)`

## 5.2.4 Execises

### 1. 
```{r 1. Had an arrival delay of two or more hours}
flights %>%
  filter(arr_delay >= 120)
```
```{r 2.Flew to Houston (IAH or HOU)}
flights %>%
  filter(dest %in% c("IAH", "HOU"))
```
```{r 3.Were operated by United, American, or Delta}
filter(flights, carrier %in% c("AA", "DL", "UA"))
```
```{r 4. Departed in summer (July, August, and September)}
filter(flights, month %in% c(7, 8, 9))
```
```{r 5.Arrived more than two hours late, but didn’t leave late}
flights %>%
  filter(arr_delay > 120 & dep_delay <= 0)
```
```{r 6.Were delayed by at least an hour, but made up over 30 minutes in flight}
flights %>%
  filter(dep_delay >= 60 & dep_delay - arr_delay > 30)
```
```{r 7.Departed between midnight and 6am (inclusive)}
flights %>%
  filter(dep_time <= 600 | dep_time == 2400)
```

### 2.Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
flights %>%
  filter(between(month, 7, 9))
```
### 3. 
```{r How many flights have a missing dep_time?}
flights %>%
  filter(is.na(dep_time)) %>%
  count()
```
```{r What other variables are missing?}
summary(flights)
## dep_delay, arr_time, arr_delay, air_time
## could be cancelled flights, potentially.
```
### 4.Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}
NA ^ 0 ## because all x^0=1
```
```{r}
NA | TRUE
```

```{r}
NA & FALSE
```
```{r}
NA * 0
```

## `arrange()`

- changes the order of rows
-  If you provide more than one column name, each additional column will be used to break ties in the values of preceding columns
```{r}
arrange(flights, year, month, day)
```

- Use desc() to re-order by a column in descending order
```{r}
arrange(flights, desc(dep_delay))
```

## `select()`

- subset with names of variables
    - `select(data, var1:var3)` subsets the first three variables
    - `select(data, -(var1:var3))` subsets everything EXCEPT the first three
    - `starts_with("abc")`
    - `ends_with("xyz")`
    - `contains("ijk")`
    - `num_range("x", 1:3)` matches x1, x2 and x3
- `select(data, var3, var7, everything())` moves var3 and va7 to the front and keeps everthing else in the data frame

### `rename()`
`rename(data, var1 = var_1)`

## `mutate()`
- `mutate(data, var = var1 * var3)`
- if only want to keep the new variables: `transmute()`

## 5.5.2 Exercises

### 1.

```{r}

```

